---
title: "Assuming 100% Voter Participation, The Liberal Party Would Win 48% of the Popular Vote in the 2019 Canadian Federal Election"
author: "Xinyi Zhang"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: |
  | In the 2019 Canadian Federal Election, incumbent Prime Minister Justin Trudeau claimed a narrow victory over the Conversative Party amid 62% voter turnout. In this paper, we develop a multi-level regression model with post-stratification by training a multinomial logistic regression model using voter survey results (from the 2019 Canadian Election Study) and predicting the outcome of the popular vote using large-scale demographic data for the Canadian population (from the 2017 General Social Survey). Our model predicts that assuming 100% voter turnout, Trudeau would have won the popular vote in the 2019 Canadian Federal Election instead of his opponent Andrew Scheer. Our breakdown of votes by demographic groups provides political parties with information on how to effectively target voters during future elections and our prediction speaks to the importance for each party to mobilize its voter base.
  |
  | **Keywords:** forecasting, 2019 Canadian Federal Election, multilevel regression with post-stratification, voter turnout
thanks: "Code and data supporting this analysis are available at: https://github.com/cindyzhang99/forecasting_canadian_election."

output:
  bookdown::pdf_document2:
toc: FALSE
bibliography: references.bib
nocite: '@*'
link-citations: yes
---

```{r setup, include=FALSE}
library(tidyverse)
```

# Introduction
In 2019, Trudeau defied election predictions and narrowly won enough seats to win reelection. This raises the question, would the outcome of the popular vote of the election have changed assuming 100% voter turnout? By choosing 5 explanatory demographic variables closely associated with voting and political affiliation (age, gender, state, race, and education), we develop a multi-level regression model with post-stratification (MRP) in order to predict the outcome of the popular vote. 

In our analysis, we modeled the relationship between our selected demographic variables and a person's vote choice (for one of the six major Canadian political parties). The comparison of the distribution of demographic variables in the individual-level Canadian Election Study dataset set (@ces) compared to the General Social Survey post-stratification dataset (@gss) revealed the importance of using multilevel regression with post-stratification because the training data is not proportional to the Canadian population. We found that we were able to make predictions with approximately 45% accuracy and discussed the ways in which the demographic attributes of each political party's voter base can be used for political strategizing. 

We conducted our data modeling and analysis using the R statistical language (@citeR) and the tidyverse (@tidyverse), janitor (@janitor), nnet (@nnet), Metrics (@metrics), cesR (@cesr), labelled (@labelled), bookdown (@bookdown_cite), and R Markdown (@rmarkdown) packages.

In Section 2, we discuss the 2 datasets we used, how they were collected and key highlights of these datasets, followed by visualizations of the data. In Section 3, we explained the construction of our model and our use of the MRP technique. Finally, we present our results in Section 4 and discuss how our results should inform future political strategies in Section 5. 

# Data
To train our model to predict voting on an individual level for the 2019 Canadian Federal Election, we used data from the 2019 Canadian Election Study (CES) (@ces). To make predictions on the outcome of the 2019 Canadian Federal Election through post-stratification, we used data from the 2017 General Social Survey (GSS) (@gss). 

In the following subsections ([Individual-level Survey Dataset](#individual-level-survey-dataset), [Post-stratification Dataset](#post-stratification-dataset)), we will discuss how each dataset was collected and highlight their key features. Then, in the [Data Visualization subsection](#data-visualization), we'll graph the distribution of our variables of interest. We will use this data in the multilevel regression with post-stratification (MRP) technique that we will describe in the [Model section](#model).

## Individual-level Survey Dataset
According to @cescodebook, from September 13, 2019 to October 21, 2019, the Consortium on Electoral Democracy conducted online surveys on political views and voting intent prior to the 2019 Canadian Federal Election. Their target population was all Canadian citizens and permanent residents, aged 18 or older. 

The online platform Qualtrics was used to conduct sampling. Qualtrics used several panels as the sampling frame for this survey. Sampling frames are lists of the individuals that will be selected for the survey sample, meaning that the members of panels aggregated by Qualtrics form a list of a subset of the target population. Then, a sample was selected from the frame using a purposive sampling method. This is a non-probability sampling method where the researcher decides which samples are most representative of the target population. According to the CES Codebook, demographic targets for province, gender (50% male and 50% female) and age (28% aged 18-34, 33% aged 35-54, and 39% aged 55 or older) were set. More specific information about the sampling method was not provided. 

Of the 74,548 individuals contacted to take the survey, about 26% did not complete the questionnaire. Another 13% exceeded demographic quotas. Lastly, approximately 10% of responses were removed for speeding (spending less than 500 seconds completing the survey) or for "straight-lining" answers (selecting the same response for all questions), resulting in a final sample size of 37,822 respondents. To ensure results were representative of the Canadian population, survey responses were weighted using data from the 2016 Census. This ensures that the discrepancy between the target population and survey responses is minimized. 

Unfortunately, the non-probability sampling method employed by Qualtrics is a major weakness of the CES surveying methodology. Although the non-response rate is relatively low, this is clearly because the sampling frame consists of individuals who on survey panels who regularly take online questionnaires. Even so, several key features of the CES methodology should be highlighted as strengths. First off, the specificity of answer choices for many questions was very detailed. In fact, when matching CES variable levels to GSS variable levels, we often had to combine levels in the CES data because the GSS data was not as specific. Additionally, the CES recontacted some participants after the election to participate in a post-election survey. Although we will not be using the post-election data, there are a multitude of applications where a longitudinal survey about the Canadian Federal Election provides valuable insights. Lastly, as we briefly touched upon when discussing the non-response rate, the CES survey results met stringent data processing standards prior to being publicly released. Therefore, we can trust that the responses represent Canadian voter sentiments to a certain extent when working with the data.

Although the non-probability sampling method is a major weakness of the CES methodology, we will ensure that our prediction of the election popular vote will be representative of the Canadian population by conducting post-stratification using data from the GSS. Therefore, we can still make valid conclusions using the CES dataset even though the responses were drawn through non-probability sampling. 

## Post-stratification Dataset
According to @gss, from February 1, 2017 to November 30, 2017, Statistics Canada gathered data on the Canadian family unit by conducting voluntary telephone interviews for Cycle 31 of the General Social Survey. Their target population was all non-institutionalized individuals living in Canada, aged 15 or older. 

Cross-sectional sampling was conducted in a two-stage design. The stratified simple random sampling method was used in the first stage. Here, the sampling frame consisted of telephone numbers from the Census grouped as households using data from Statistic Canada's dwelling frame. Strata were formed at the census metropolitan area (CMA) level and at the province level (i.e., large CMAs formed their own stratum, smaller CMAs were grouped together, and the non-CMA regions of each province were grouped together), forming a total of 27 non-overlapping strata. Finally, households were sampled randomly from each stratum such that the number sampled units from each stratum corresponded to the population sizes of each stratum. To reiterate, the sampled population for this first stage was the chosen households from each stratum. The stratified simple random sampling method was also used in the second stage. Here, the sampling frame was a list of household members, aged 15 and older, from the households selected in the first stage. Then, one individual was randomly selected from each household, forming the sampled population. Approximately 43,000 individuals were contacted to participate in the survey.

Statistics Canada reported that the non-response rate was 52.4%. This presents problems for data analysis based on survey data if respondents differ significantly from non-respondents. To reduce the effects of non-response bias, survey estimates were adjusted based on the demographic characteristics of households that were non-responsive (by pulling their information from the 2016 Census). Another source of non-sampling error is imperfect coverage. For example, households without telephones are excluded from the sampling frame. Again, survey estimates were adjusted by weighing responses to represent all individuals in the target population. Lastly, another weakness of the survey methodology is the exclusion of the Canadian population residing in the Northwest Territories, Nunavut, and the Yukon Territory. As we found when trying to match variable levels in the CES dataset to the GSS dataset, we had to drop responses from the territories in the survey data because of the lack of information available for the territories in the post-stratification data. Due to this limitation in the GSS data coverage, our prediction of the popular vote of the 2019 Canadian Federal Election excludes the voters in the territories of Canada.

On the other hand, several key features of the GSS stand out as strengths of their surveying methodology. A major strength of the questionnaire is that it contains focused questions that comprehensively and extensively capture the subject of interest (the Canadian family). Extensive research and testing was conducted when designing the questionnaire. Upon reading through the questionnaire made available by Statistics Canada, the wording of each question is precise and clear, leaving little room for ambiguity. Additionally, another strength of the survey is that a vast majority of questions were objective (dates, events, counts) removing potential response biases that occur with subjective questions. 

Overall, the GSS surveying method using two-stage simple random stratified sampling is effective in generating a sample that is geographically representative of the population living in the Canadian provinces. Despite limitations in sampling coverage, the demographic information available is much more detailed than in comparable post-stratification datasets (e.g., the 2016 Census).

## Data Visualization
The full 2019 CES Online Survey and 2017 GSS datasets contain tens of thousands of observations for over 400 variables. 

In the interest of space, we will only discuss the variables in the two datasets that are relevant to our multinomial logistic regression model. We chose five demographic variables to serve as a independent factors that determine vote choice. These variables were present in both datasets and typically associated with political affiliation. They are:

* age (represented by the variables named `cps19_yob` in the CES dataset and `agec` in the GSS dataset): the age of the respondent at the time of the survey
* sex (`cps19_gender` in CES and `sex` in GSS): sex of the respondent
* province (`cps19_province` in CES and `prv` in GSS): province the respondent resides in
* education (`cps19_education` in CES and `ehg3_01b` in GSS): highest education level of the respondent
* religion importance (`cps19_rel_imp` in CES `rlr_110` in GSS): how important religion is to the respondent

To prepare the datasets, we selected responses from individuals who were 18 years of age or older. Additionally, we mapped the integer values of the `age` variable to 6 bins ("18 to 29", "30 to 41", "42 to 53", "54 to 65", "66 to 77", and "78 and above"). Comparing the two distributions, notice that individuals aged 78 and above are underrepresented in the CES dataset (Figure \@ref(fig:agedist)). This is likely because Canadians that are much older rarely serve as participants on online survey panels. This speaks to the importance of using post-stratification to adjust our predictions to be more representative of the Canadian population.
```{r agedist, fig.cap="Distribution of Age in CES and GSS datasets.", echo=FALSE, message=FALSE,  fig.show="hold", out.width="50%"}
ces <- read_csv("../../inputs/ces/ces_input.csv")

# find percentage of each age range
age_distribution <- ces %>% dplyr::count(age) %>% mutate(Percent = n/nrow(ces))
# plot distribution as bar graph
ggplot(data=age_distribution, aes(x=age, y=Percent, fill=Percent)) +
  geom_bar(stat="identity") +
  labs(x="Age", y="Percent", title = "Distribution of Age in CES") +
  theme(plot.title = element_text(hjust = 0.5, color = "#666666")) +
  geom_text(aes(label=paste0(round(Percent*100), "%")), position=position_dodge(width=0.9), vjust=-0.25)

gss <- read_csv("../../inputs/gss/gss_input.csv")

# find percentage of each age range
age_distribution <- gss %>% dplyr::count(age) %>% mutate(Percent = n/nrow(gss))
# plot distribution as bar graph
ggplot(data=age_distribution, aes(x=age, y=Percent, fill=Percent)) +
  geom_bar(stat="identity") +
  labs(x="Age", y="Percent", title = "Distribution of Age in GSS") +
  theme(plot.title = element_text(hjust = 0.5, color = "#666666")) +
  geom_text(aes(label=paste0(round(Percent*100), "%")), position=position_dodge(width=0.9), vjust=-0.25)
```
In the CES dataset, we dropped responses of "Other" for the variable `sex` because the GSS dataset had only "Male" and "Female" levels. Comparing the distributions of sex in each survey, females are clearly overrepresented in the CES dataset (Figure \@ref(fig:sexdist)).
```{r sexdist, fig.cap="Distribution of Sex in CES and GSS datasets.", echo=FALSE, message=FALSE, fig.show="hold", out.width="50%"}
# find percentage of each gender
gender_distribution <- ces %>% dplyr::count(cps19_gender) %>% mutate(percent = n/nrow(ces))
# plot distribution as pie chart
ggplot(gender_distribution, aes(x="", y=percent, fill=cps19_gender)) + geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) + geom_text(aes(label = paste0(round(percent*100), "%")), position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values=c("#f7b6d2", "#9edae5")) + 
  labs(x = NULL, y = NULL, fill = NULL, title = "Distribution of Sex in CES") +
  theme_classic() + theme(axis.line = element_blank(),
                          axis.text = element_blank(),
                          axis.ticks = element_blank(),
                          plot.title = element_text(hjust = 0.5, color = "#666666"))

# find percentage of each gender
gender_distribution <- gss %>% count(sex) %>% mutate(percent = n/nrow(gss))
# plot distribution as pie chart
ggplot(gender_distribution, aes(x="", y=percent, fill=sex)) + geom_bar(stat="identity", width=1) + 
  coord_polar("y", start=0) + geom_text(aes(label = paste0(round(percent*100), "%")), position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values=c("#f7b6d2", "#9edae5")) + 
  labs(x = NULL, y = NULL, fill = NULL, title = "Distribution of Sex in GSS") +
  theme_classic() + theme(axis.line = element_blank(),
                          axis.text = element_blank(),
                          axis.ticks = element_blank(),
                          plot.title = element_text(hjust = 0.5, color = "#666666"))
```
In the CES dataset, we dropped responses of "Northwest Territories", "Nunavut", and "Yukon" for the variable `province` because the GSS dataset lacked levels for the territories. Comparing the distributions of provinces in each survey, all of the other provinces are underrepresented while Ontario is overrepresented (by 50%) (Figure \@ref(fig:provdist)). Clearly, this has significant implications for our prediction of the popular vote because if we were just to use the results of the CES dataset, the opinions of Canadians residing in Ontario would be significantly overrepresented. Luckily, we are conducting post-stratification using the GSS dataset, which better represents the demographics of the Canadian population.
```{r provdist, fig.cap="Distribution of Provinces in CES and GSS datasets.", echo=FALSE, message=FALSE,  fig.show="hold", out.width="50%"}
# find percentage of each province
province_distribution <- ces %>% dplyr::count(cps19_province) %>% mutate(Percent = n/nrow(ces))
# plot distribution as bar graph
ggplot(data=province_distribution, aes(x=cps19_province, y=Percent, fill=Percent)) +
  geom_bar(stat="identity") +
  labs(x="Province", y="Percent", title = "Distribution of Provinces in CES") +
  theme(axis.text.x = element_text(angle = 25), 
        plot.title = element_text(hjust = 0.5, color = "#666666")) +
  geom_text(aes(label=paste0(round(Percent*100), "%")), position=position_dodge(width=0.9), vjust=-0.25)

# find percentage of each province
province_distribution <- gss %>% dplyr::count(province) %>% mutate(Percent = n/nrow(gss))
# plot distribution as bar graph
ggplot(data=province_distribution, aes(x=province, y=Percent, fill=Percent)) +
  geom_bar(stat="identity") +
  labs(x="Province", y="Percent", title = "Distribution of Provinces in GSS") +
  theme(axis.text.x = element_text(angle = 25), 
        plot.title = element_text(hjust = 0.5, color = "#666666")) +
  geom_text(aes(label=paste0(round(Percent*100), "%")), position=position_dodge(width=0.9), vjust=-0.25)
```
We simplified the levels for the variable `education` in both datasets, highlighting only the most important categories ("High school or less", "Some college or trade school", "Undergraduate degree", "Some graduate school or more"). Individuals who pursued any level of post-secondary education are overrepresented and individuals who have at most a high school diploma are underrepresented in the CES dataset (Figure \@ref(fig:edudist)). Again, this has significant implications for our prediction of the popular vote if we were to use only the CES dataset but our post-stratification technique using the GSS dataset will mitigate this discrepancy between survey respondents and the demographics of the Canadian population.
```{r edudist, fig.cap="Distribution of Education in CES and GSS datasets.", echo=FALSE, message=FALSE,  fig.show="hold", out.width="50%"}

# find percentage of each education level
education_distribution <- ces %>% dplyr::count(cps19_education) %>% mutate(percent = n/nrow(ces))
# reorder levels
education_distribution$cps19_education <- factor(education_distribution$cps19_education, 
                                                 levels=c("High school or less",
                                                          "Some college or trade school",
                                                          "Undergraduate degree",
                                                          "Some graduate school or more"))
# plot distribution as pie chart
ggplot(education_distribution, aes(x="", y=percent, fill=cps19_education)) + geom_bar(stat="identity", width=1) + 
  coord_polar("y", start=0) + geom_text(aes(label = paste0(round(percent*100), "%")), position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values=c("#b1d8b7", "#76b947", "#54902a", "#2f5233")) +
  labs(x = NULL, y = NULL, fill = NULL, title = "Distribution of Education in CES") + 
  theme_classic() + theme(axis.line = element_blank(),
                          axis.text = element_blank(),
                          axis.ticks = element_blank(),
                          plot.title = element_text(hjust = 0.5, color = "#666666"))

# find percentage of each education level
education_distribution <- gss %>% dplyr::count(education) %>% mutate(percent = n/nrow(gss))
# reorder levels
education_distribution$education <- factor(education_distribution$education, 
                                           levels=c("High school or less",
                                                    "Some college or trade school",
                                                    "Undergraduate degree",
                                                    "Some graduate school or more"))

# plot distribution as pie chart
ggplot(education_distribution, aes(x="", y=percent, fill=education)) + geom_bar(stat="identity", width=1) + 
  coord_polar("y", start=0) + geom_text(aes(label = paste0(round(percent*100), "%")), position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values=c("#b1d8b7", "#76b947", "#54902a", "#2f5233")) +
  labs(x = NULL, y = NULL, fill = NULL, title = "Distribution of Education in GSS") + 
  theme_classic() + theme(axis.line = element_blank(),
                          axis.text = element_blank(),
                          axis.ticks = element_blank(),
                          plot.title = element_text(hjust = 0.5, color = "#666666"))
```
Luckily, the levels for religion importance perfectly overlapped in both datasets. However, the distribution of each level in both datasets indicates that individuals with strongly held religious opinions (religion is "Very important" or "Not at all important") are underrepresented while individuals with middle of the road opinions ("Somewhat important", "Not very important") are overrepresented in the CES dataset (Figure \@ref(fig:religiondist)). Once again, we will be using the GSS dataset for post-stratification to ensure our prediction is more representative of the Canadian population than the demographics of the respondents to the CES is.
```{r religiondist, fig.cap="Distribution of Religion Importance in CES and GSS datasets.", echo=FALSE, message=FALSE,  fig.show="hold", out.width="50%"}
# find percentage of each religion level
religion_distribution <- ces %>% dplyr::count(cps19_rel_imp) %>% mutate(percent = n/nrow(ces))
# plot distribution as pie chart
ggplot(religion_distribution, aes(x="", y=percent, fill=cps19_rel_imp)) + geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) + geom_text(aes(label = paste0(round(percent*100), "%")), position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values=c("#b1d8b7", "#94c973", "#76b947", "#54902a", "#2f5233")) +
  labs(x = NULL, y = NULL, fill = NULL, title = "Distribution of Religion Importance in CES") + 
  theme_classic() + theme(axis.line = element_blank(),
                          axis.text = element_blank(),
                          axis.ticks = element_blank(),
                          plot.title = element_text(hjust = 0.5, color = "#666666"))

# find percentage of each religion level
religion_distribution <- gss %>% dplyr::count(religion_importance) %>% mutate(percent = n/nrow(gss))
# plot distribution as pie chart
ggplot(religion_distribution, aes(x="", y=percent, fill=religion_importance)) + geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) + geom_text(aes(label = paste0(round(percent*100), "%")), position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values=c("#b1d8b7", "#94c973", "#76b947", "#54902a", "#2f5233")) +
  labs(x = NULL, y = NULL, fill = NULL, title = "Distribution of Religion Importance in GSS") + 
  theme_classic() + theme(axis.line = element_blank(),
                          axis.text = element_blank(),
                          axis.ticks = element_blank(),
                          plot.title = element_text(hjust = 0.5, color = "#666666"))
```
\newpage

# Model
The purpose of our model is to predict the popular vote of the 2019 Canadian Federal Election assuming 100% voter turnout. We found when exploring and graphing the data in the in the [Data Visualization section](#data-visualization), the CES has underrepresented or overrepresented certain demographics compared to their actual distributions in the population. Namely, older voters are underrepresented, women are overrepresented, voters from Ontario are overrepresented, voters with some form of post-secondary education are overrepresented, and so forth. Therefore, if we were to predict the outcome of the popular vote based on the demographic proportions found in the CES dataset, we would find that the prediction is biased because the survey sample is not representative of the Canadian population. Hence, we must use multilevel regression with post-stratification (MRP) to base our estimate of the popular vote on demographic distributions are are closer to the actual population distributions. This method can be broken down into two parts: multinomial regression modeling and post-stratification.

First, we want to train a model on the CES dataset to predict vote choice. Our dependent variable `vote` is a categorical variable with six possible values ("New Democratic Party", "Green Party", "Liberal Party", "Conservative Party", "People's Party", and "Bloc Québécois"), which describes how a respondent would vote in the 2019 Canadian Federal Election. The independent variables in our model are the demographic characteristics of the respondent (age, gender, province, education, and religion importance). (As previously established, these were chosen based on their association with determining political affiliation and their persistence in both datasets).

Since we are representing the vote as a classification problem (with more than 2 possible values), multinomial logistic regression is the most suitable choice to model the relationship between vote choice and our independent demographic variables. Given an input of values for the demographic variables, the model will output six probabilities, one for each possible vote choice. The vote choice corresponding to the largest probability is our model's prediction for the given demographic values.

We fully derive the mathematical representation of the multinomial logistic regression model in the [Appendix](#derivation-of-multinomial-logistic-regression-model). In the interest of space, here we will just provide the equations we derived for the probability of each outcome:

$$ \Pr(Y_i = 1) = \frac{e^{\boldsymbol{\beta_1 \cdot X_i}}}{1 + \sum_{k\in \{1,2,3,5,6\}} e^{\boldsymbol{\beta_k \cdot X_i}}} $$
$$ \Pr(Y_i = 2) = \frac{e^{\boldsymbol{\beta_2 \cdot X_i}}}{1 + \sum_{k\in \{1,2,3,5,6\}} e^{\boldsymbol{\beta_k \cdot X_i}}} $$
$$ \Pr(Y_i = 3) = \frac{e^{\boldsymbol{\beta_3 \cdot X_i}}}{1 + \sum_{k\in \{1,2,3,5,6\}} e^{\boldsymbol{\beta_k \cdot X_i}}} $$
$$ \Pr(Y_i = 4) = \frac{1}{1 + \sum_{k\in \{1,2,3,5,6\}} e^{\boldsymbol{\beta_k \cdot X_i}}}$$
$$ \Pr(Y_i = 5) = \frac{e^{\boldsymbol{\beta_5 \cdot X_i}}}{1 + \sum_{k\in \{1,2,3,5,6\}} e^{\boldsymbol{\beta_k \cdot X_i}}} $$
$$ \Pr(Y_i = 6) = \frac{e^{\boldsymbol{\beta_6 \cdot X_i}}}{1 + \sum_{k\in \{1,2,3,5,6\}} e^{\boldsymbol{\beta_k \cdot X_i}}} $$
We will explain each variable that appears in the equations above (again, please see the [Appendix](#derivation-of-multinomial-logistic-regression-model) for a more natural definition of each variable as we derive these equations). 
$Y_i$ represents the outcome of the response variable for the $i$th observation. Since we have a total of 6 possible outcomes ("New Democratic Party", "Green Party", "Liberal Party", "Conservative Party", "People's Party", and "Bloc Québécois"), we represent each of them as 1, 2, 3, 4, 5, 6 in these equations in the interest of space. 

$\boldsymbol{\beta_k}$ is the row vector with elements that are the coefficients for the explanatory variables for the $k$th outcome. More explicitly, $\boldsymbol{\beta_k} = [\beta_{0,k}, \beta_{1,k}, \beta_{2,k}, \beta_{3,k}, \beta_{4,k}, \beta_{5,k}]$ where $\beta_{0,k}$ is the intercept for the $k$th outcome, $\beta_{1,k}$ is the coefficient for the first explanatory variable (age) for the $k$th outcome, $\beta_{2,k}$ is the coefficient for the second explanatory variable (sex) for the $k$th outcome, $\beta_{3,k}$ is the coefficient for the third explanatory variable (province) for the $k$th outcome, $\beta_{4,k}$ is the coefficient for the fourth explanatory variable (education) for the $k$th outcome, and $\beta_{5,k}$ is the coefficient for the fifth explanatory variable (religion importance) for the $k$th outcome.

$\boldsymbol{x_i}$ is the row vector of values for the explanatory variables for the $i$th observation. Specifically, $\boldsymbol{x_i} = [1, x_{1,i}, x_{2,i}, x_{3,i}, x_{4,i}, x_{5,i}]$ where $x_{1,i}$ is the value of the first explanatory variable (age) for the $i$th observation, $x_{2, i}$ is the value of the second explanatory variable (sex) for the $i$th observation, and so on.

Therefore, the first equation represents the probability that the vote choice of the $i$th observation is "New Democratic Party" given values for age, sex, province, education, and religion importance The other 5 equations can be interpreted in a similar manner.

We implemented the multinomial logistic regression model using `multinom()` from the nnet library (@nnet). At first the model failed to converge because the maximum number of iterations would be reached before convergence. Consequently, we increased the maximum number of iterations from the default value of 100 to 200 and we observed the model converging after approximately 120 iterations.

We tested 3 different models using cross-validation testing. Assuming that age, gender, and province are significant predictors of vote choice, we tested different combinations of education level and religion importance. Model 1 has age, gender, province, and education level as independent variables. Model 2 has age, gender, province, and religion importance as independent variables. Lastly, model 3 has all 5 demographic variables as independent variables (age, gender, province, education level, and religion importance). To conduction cross-validation, we shuffled the dataset into training and testing subsets (an 80-20 split, respectively). Then, we trained the model using only the training data and made predictions on the testing data. Finally, we compared the model's predictions to the ground truth values for vote choice for the test set. Model 1's predictions for the test set had an accuracy of 44.0% while model 2 and model 3 had accuracies of 44.2% and 45.1%, respectively. Therefore, we proceeded with the use of model 3 for making predictions on the post-stratification dataset because it had the highest accuracy out of the models we implemented.

Finally, we used post-stratification to accurately predict the popular vote because of the discrepancy between the demographic distribution of the CES respondents and the demographics of the Canadian  population. We conducted post-stratification by making predictions on all combinations of explanatory variables in the GSS. 

# Results
In Figure \@ref(fig:votedist), we see that the distribution of votes in the CES is in favor of the Conservative party, followed closely by the Liberal Party. However, as we've noticed when observing the distribution of our chosen demographic factors, the CES dataset does not appear to be representative of the Canadian population. Therefore, it is imperative that we use post-stratification when predicting the popular vote of the 2019 Canadian Federal Election to mitigate the effects of sampling bias.

```{r votedist, echo=FALSE, fig.cap="Distribution of Votes in CES.", message=FALSE, fig.align="center", fig.height=3}
# find percentage of each vote
vote_distribution <- ces %>% dplyr::count(cps19_vote) %>% mutate(Percent = n/nrow(ces))
# reorder levels
vote_distribution$cps19_vote <- factor(vote_distribution$cps19_vote, 
                                       levels=c("New Democratic Party",
                                                "Green Party",
                                                "Liberal Party",
                                                "Conservative Party",
                                                "People's Party",
                                                "Bloc Québécois"))
# plot distribution as bar graph
ggplot(data=vote_distribution, aes(x=cps19_vote, y=Percent, fill=Percent)) +
  geom_bar(stat="identity") +
  labs(x="Province", y="Percent") +
  theme(axis.text.x = element_text(angle = 25)) +
  geom_text(aes(label=paste0(round(Percent*100), "%")), position=position_dodge(width=0.9), vjust=-0.25)
```

In Figure \@ref(fig:poststratdist), we see that the distribution of votes in post-stratification is led by the Liberal Party with 48% of the vote followed by Conservative Party with 44% of the vote. 

```{r poststratdist, echo=FALSE, fig.cap="Distribution of Votes in Post-Stratification.", message=FALSE, fig.align="center", fig.height=3}
poststrat <- read_csv("../post_strat.csv")
counttb <- count(x=poststrat, vote, wt=freq)
total <- sum(counttb$n)
counttb$Proportion <- counttb$n/total

# plot distribution as bar graph
ggplot(data=counttb, aes(x=vote, y=Proportion, fill=Proportion)) +
  geom_bar(stat="identity") +
  labs(x="Political Party", y="Proportion") +
  theme(axis.text.x = element_text(angle = 25)) +
  geom_text(aes(label=paste0(round(Proportion*100), "%")), position=position_dodge(width=0.9), vjust=-0.25)
```

Comparing voting choice broken down by age group in the CES data (Figure \@ref(fig:cesagevote)) and after post-stratification (Figure \@ref(fig:postagevote)), both graphs (to varying degrees) indicate that as age increases, an increasing proportion of voters vote for conservative parties. 

Comparing voting choice broken down by gender in the CES data (Figure \@ref(fig:cesgendervote)) and after post-stratification (Figure \@ref(fig:postgendervote)), both graphs (to varying degrees) indicate that females are more likely to vote for more liberal parties than males. 

Comparing voting choice broken down by province in the CES data (Figure \@ref(fig:cesprovincevote)) and after post-stratification (Figure \@ref(fig:postprovincevote)), we can see that the post-stratification graph contrasts the provinces that lean conservative (Alberta, Saskatchewan, Manitoba) with the provinces that lean liberal (Newfoundland and Labrador, Nova Scotia, Prince Edward Island), while provinces such as New Brunswick and Ontario are split almost evenly. 

As education level increases, so do the proportion of voters who would vote for the Liberal Party (Figures \@ref(fig:ceseducationvote) and \@ref(fig:posteducationvote)). Interestingly, both graphs indicate that as education increases, the proportion of voters who would vote for the New Democratic Party slightly decreases (likely related to the trend observed with regards to age).

A curvilinear relationship between religion importance and vote choice can be observed in Figures \@ref(fig:cesreligionvote) and \@ref(fig:postreligionvote). Ignoring the "Don't know" factor, individuals who place more value on religion are more likely to vote conservatively while individuals who find religion unimportant are more likely to vote liberally.

\newpage 
```{r cesagevote, fig.cap="Distribution of Votes by Age Group in CES.", fig.height=3, echo=FALSE, message=FALSE}
# load original cleaned ces data
ces <- read_csv("../../inputs/ces/ces_input.csv")
# load post-stratification data
poststrat <- read_csv("../../outputs/post_strat.csv")

# set up custom color scheme where blue indicates the more liberal parties and red indicates the more conservative parties
# match colors across ces and post-strat
colors <- c("Green Party" = "#2166AC", "New Democratic Party" = "#4393C3", "Liberal Party" = "#92C5DE",
            "Conservative Party" = "#F4A582", "People's Party" = "#D6604D", "Bloc Québécois" = "#B2182B")

# find percentage of each vote by age groups in ces
age_distribution_ces <- count(ces, age, cps19_vote) %>% group_by(age) %>% mutate(Percent = n/sum(n))
# reorder levels
age_distribution_ces$cps19_vote <- factor(age_distribution_ces$cps19_vote, 
                                       levels=c("Green Party",
                                                "New Democratic Party",
                                                "Liberal Party",
                                                "Conservative Party",
                                                "People's Party",
                                                "Bloc Québécois"))
# plot distribution with respect to age as line graph
ggplot(age_distribution_ces, aes(x=age, y=Percent, group=cps19_vote, color=cps19_vote)) +
  geom_point() + 
  geom_line() + 
  labs(x="Age", y="Vote Proportion", color="Political Party") + 
  scale_color_manual(values=colors)
```

```{r postagevote, fig.cap="Distribution of Votes by Age Group in Post-Stratification.", fig.height=3, echo=FALSE, message=FALSE}
# find percentage of each vote by age groups in poststrat
age_distribution_poststrat <- aggregate(freq ~ age + vote, data=poststrat, sum) %>% group_by(age) %>% mutate(Percent = freq/sum(freq))
# reorder levels
age_distribution_ces$cps19_vote <- factor(age_distribution_ces$cps19_vote, 
                                          levels=c("Green Party",
                                                   "New Democratic Party",
                                                   "Liberal Party",
                                                   "Conservative Party",
                                                   "People's Party",
                                                   "Bloc Québécois"))
# plot distribution with respect to age as line graph
ggplot(age_distribution_poststrat, aes(x=age, y=Percent, group=vote, color=vote)) +
  geom_point() + 
  geom_line() + 
  labs(x="Age", y="Vote Proportion", color="Political Party") +
  scale_color_manual(values=colors)
```


```{r cesgendervote, fig.cap="Distribution of Votes by Sex in CES.", fig.height=4, echo=FALSE, message=FALSE}
# find percentage of each vote by sex in ces
sex_distribution_ces <- count(ces, cps19_gender, cps19_vote) %>% group_by(cps19_gender) %>% mutate(Percent = n/sum(n))
# reorder levels
sex_distribution_ces$cps19_vote <- factor(sex_distribution_ces$cps19_vote, 
                                          levels=c("Green Party",
                                                   "New Democratic Party",
                                                   "Liberal Party",
                                                   "Conservative Party",
                                                   "People's Party",
                                                   "Bloc Québécois"))

# plot distribution with respect to sex as line graph
ggplot(sex_distribution_ces, aes(x=cps19_gender, y=Percent, group=cps19_vote, color=cps19_vote)) +
  geom_point() + 
  geom_line() + 
  labs(x="Sex", y="Vote Proportion", color="Political Party") + 
  scale_color_manual(values=colors)
```

```{r postgendervote, fig.cap="Distribution of Votes by Sex in Post-Stratification.", fig.height=4, echo=FALSE, message=FALSE}
# find percentage of each vote by sex groups in poststrat
sex_distribution_poststrat <- aggregate(freq ~ sex + vote, data=poststrat, sum) %>% group_by(sex) %>% mutate(Percent = freq/sum(freq))
# reorder levels
sex_distribution_poststrat$vote <- factor(sex_distribution_poststrat$vote, 
                                          levels=c("Green Party",
                                                   "New Democratic Party",
                                                   "Liberal Party",
                                                   "Conservative Party",
                                                   "People's Party",
                                                   "Bloc Québécois"))
# plot distribution with respect to sex as line graph
ggplot(sex_distribution_poststrat, aes(x=sex, y=Percent, group=vote, color=vote)) +
  geom_point() + 
  geom_line() + 
  labs(x="Sex", y="Vote Proportion", color="Political Party") +
  scale_color_manual(values=colors)
```


```{r cesprovincevote, fig.cap="Distribution of Votes by Province in CES.", fig.height=4, echo=FALSE, message=FALSE}
# find percentage of each vote by province in ces
province_distribution_ces <- count(ces, cps19_province, cps19_vote) %>% group_by(cps19_province) %>% mutate(Percent = n/sum(n))
# reorder levels
province_distribution_ces$cps19_vote <- factor(province_distribution_ces$cps19_vote, 
                                          levels=c("Green Party",
                                                   "New Democratic Party",
                                                   "Liberal Party",
                                                   "Conservative Party",
                                                   "People's Party",
                                                   "Bloc Québécois"))

# plot distribution with respect to province as line graph
ggplot(province_distribution_ces, aes(x=cps19_province, y=Percent, group=cps19_vote, color=cps19_vote)) +
  geom_point() + 
  geom_line() + 
  labs(x="Province", y="Vote Proportion", color="Political Party") + 
  theme(axis.text.x = element_text(angle = 25)) +
  scale_color_manual(values=colors)
```

```{r postprovincevote, fig.cap="Distribution of Votes by Province in Post-Stratification.", fig.height=4, echo=FALSE, message=FALSE}
# find percentage of each vote by province in poststrat
province_distribution_poststrat <- aggregate(freq ~ province + vote, data=poststrat, sum) %>% group_by(province) %>% mutate(Percent = freq/sum(freq))
# reorder levels
province_distribution_poststrat$vote <- factor(province_distribution_poststrat$vote, 
                                          levels=c("Green Party",
                                                   "New Democratic Party",
                                                   "Liberal Party",
                                                   "Conservative Party",
                                                   "People's Party",
                                                   "Bloc Québécois"))
# plot distribution with respect to province as line graph
ggplot(province_distribution_poststrat, aes(x=province, y=Percent, group=vote, color=vote)) +
  geom_point() + 
  geom_line() + 
  labs(x="Province", y="Vote Proportion", color="Political Party") +
  theme(axis.text.x = element_text(angle = 25)) +
  scale_color_manual(values=colors)
```


```{r ceseducationvote, fig.cap="Distribution of Votes by Education in CES.", fig.height=4, echo=FALSE, message=FALSE}
# find percentage of each vote by education in ces
education_distribution_ces <- count(ces, cps19_education, cps19_vote) %>% group_by(cps19_education) %>% mutate(Percent = n/sum(n))
# reorder levels
education_distribution_ces$cps19_vote <- factor(education_distribution_ces$cps19_vote, 
                                               levels=c("Green Party",
                                                        "New Democratic Party",
                                                        "Liberal Party",
                                                        "Conservative Party",
                                                        "People's Party",
                                                        "Bloc Québécois"))
# reorder education levels
education_distribution_ces$cps19_education <- factor(education_distribution_ces$cps19_education, 
                                                levels=c("High school or less",
                                                         "Some college or trade school",
                                                         "Undergraduate degree",
                                                         "Some graduate school or more"))
# plot distribution with respect to education as line graph
ggplot(education_distribution_ces, aes(x=cps19_education, y=Percent, group=cps19_vote, color=cps19_vote)) +
  geom_point() + 
  geom_line() + 
  labs(x="Education", y="Vote Proportion", color="Political Party") + 
  theme(axis.text.x = element_text(angle = 10)) +
  scale_color_manual(values=colors)
```

```{r posteducationvote, fig.cap="Distribution of Votes by Education in Post-Stratification.", fig.height=4, echo=FALSE, message=FALSE}
# find percentage of each vote by education in poststrat
education_distribution_poststrat <- aggregate(freq ~ education + vote, data=poststrat, sum) %>% group_by(education) %>% mutate(Percent = freq/sum(freq))
# reorder levels
education_distribution_poststrat$vote <- factor(education_distribution_poststrat$vote, 
                                               levels=c("Green Party",
                                                        "New Democratic Party",
                                                        "Liberal Party",
                                                        "Conservative Party",
                                                        "People's Party",
                                                        "Bloc Québécois"))
# reorder education levels
education_distribution_poststrat$education <- factor(education_distribution_poststrat$education, 
                                                     levels=c("High school or less",
                                                              "Some college or trade school",
                                                              "Undergraduate degree",
                                                              "Some graduate school or more"))
# plot distribution with respect to education as line graph
ggplot(education_distribution_poststrat, aes(x=education, y=Percent, group=vote, color=vote)) +
  geom_point() + 
  geom_line() + 
  labs(x="Education", y="Vote Proportion", color="Political Party") +
  theme(axis.text.x = element_text(angle = 10)) +
  scale_color_manual(values=colors)
```


```{r cesreligionvote, fig.cap="Distribution of Votes by Education in CES.", fig.height=4, echo=FALSE, message=FALSE}
religion_distribution_ces <- count(ces, cps19_rel_imp, cps19_vote) %>% group_by(cps19_rel_imp) %>% mutate(Percent = n/sum(n))
# reorder levels
religion_distribution_ces$cps19_vote <- factor(religion_distribution_ces$cps19_vote, 
                                                levels=c("Green Party",
                                                         "New Democratic Party",
                                                         "Liberal Party",
                                                         "Conservative Party",
                                                         "People's Party",
                                                         "Bloc Québécois"))
# plot distribution with respect to religion importance as line graph
ggplot(religion_distribution_ces, aes(x=cps19_rel_imp, y=Percent, group=cps19_vote, color=cps19_vote)) +
  geom_point() + 
  geom_line() + 
  labs(x="Religion Importance", y="Vote Proportion", color="Political Party") + 
  theme(axis.text.x = element_text(angle = 10)) +
  scale_color_manual(values=colors)
```

```{r postreligionvote, fig.cap="Distribution of Votes by Education in Post-Stratification.", fig.height=4, echo=FALSE, message=FALSE}
# find percentage of each vote by religion importance in poststrat
religion_distribution_poststrat <- aggregate(freq ~ religion_importance + vote, data=poststrat, sum) %>% group_by(religion_importance) %>% mutate(Percent = freq/sum(freq))
# reorder levels
religion_distribution_poststrat$vote <- factor(religion_distribution_poststrat$vote, 
                                                levels=c("Green Party",
                                                         "New Democratic Party",
                                                         "Liberal Party",
                                                         "Conservative Party",
                                                         "People's Party",
                                                         "Bloc Québécois"))
# plot distribution with respect to religion importance as line graph
ggplot(religion_distribution_poststrat, aes(x=religion_importance, y=Percent, group=vote, color=vote)) +
  geom_point() + 
  geom_line() + 
  labs(x="Religion Importance", y="Vote Proportion", color="Political Party") +
  theme(axis.text.x = element_text(angle = 10)) +
  scale_color_manual(values=colors)
```

\newpage
# Discussion, Weaknesses, and Next Steps
From our visualization of the CES dataset, we determined that the demographic distribution of the survey did not correspond to that of the population. Consequently, upon training a multinomial logistic model on the CES dataset to predict vote choice, we conducted post-stratification using the GSS dataset which better captures the demographic distribution of Canadian voters. 

However, our predicted results for the popular vote assuming 100% voter turnout (Figure \@ref(fig:poststratdist)) does not seem very realistic with the Liberal Party expected to win with 48% of the popular vote followed by the Conservative Party with 44%. The Bloc Québécois are expected to win 4% of the popular vote followed by 3% for the New Democratic Party and close to 0% for the Green Party. We also predict that the the People's Party would receive exactly 0% of the popular vote. Comparing these results to the actual results of the popular vote in the 2019 Canadian Federal Election, it seems that predictions were overwhelmingly in favor of the Liberal and Conservative Parties while votes for the other political parties were underpredicted.

Considering our methodology, the cause of this interesting result is likely two-fold. 

Firstly, we determined through cross-validation that our model has an accuracy of about 45% when predicting vote choice given five demographic attributes (age, sex, province, education level, and religion importance). Clearly, this result is much better than chance given that there are six possible options. Furthermore, being able to predict someone's vote with a certain degree of accuracy given only 5 demographic variables is fairly impressive considering political affiliation does not follow any hard and fast rules. However, this low accuracy is likely one of the sources of error when it comes to our unrealistic popular vote prediction. Our model likely found it difficult to distinguish between the New Democratic Party, the Green Party, and the Liberal Party when predicting the vote of an individual with liberal-leaning demographic attributes (and likewise with regards to the People's Party and the Conservative Party when predicting the vote of an individual with conservative-leaning demographic attributes). In fact, we chose a multinomial logistic regression model because we wanted to be able to predict the popular vote realistically given the choice of all six major political parties but an unfortunate side effect of this choice is that datapoints concerning smaller political parties were overshadowed by those concerning the larger political parties (Liberal Party and Conservative Party). Consequently, the model likely will predict "Liberal Party" or "Conservative Party" over the smaller political parties because the prediction is more likely to be correct, simply because there are more Liberal Party and Conservative Party vote choice data points.

The second cause of our unrealistic popular vote prediction is the fact that not every combination of demographic variables was available in the GSS. Therefore, the combinations of attributes that are missing from post-stratification are simply not represented in our calculation of the popular vote. This would also cause our unrealistic popular vote prediction if some demographics that would result in a vote prediction of one of the smaller parties were not represented in the post-stratification dataset. 

Overall, we can trust that our results indicate more popular votes would go to liberal political parties than conservative. This is because demographic variables associated with more liberal-leaning individuals (such as being younger, being female, having some form of postsecondary education, and placing little to no importance on religion) are more prevalent in the Canadian population.  However, due to limitations in the survey data we used to train our model, limitations with regards to the accuracy of the model itself, and limitations with regards to the post-stratification dataset we used to predict the popular vote, it seems that our prediction of the popular vote assuming 100% voter turnout with regards to all six political parties is inaccurate.

Even so, our results are significant because they indicate which factors are most important in determining vote choice as well as the demographics of each political party. For example, the majority of the New Democratic Party voter base are likely younger and/or female Canadians who find religion unimportant. The majority of the Liberal Party voter base are younger, highly educated, and/or female Canadians who find religion unimportant. On the other hand, Conversative Party voters are older and/or male Canadians who find religion very important and/or have at most a high school education. 

Although there are more liberal voters than conservative voters, the Conservative Party won the majority of the popular vote in the 2019 Canadian Federal Election and narrowly lost the seat of the prime minister to the Liberal Party (who won more seats despite coming in second in the popular vote). To continue to win the most seats and to win the popular vote in the 2023 Canadian Federal Election, the Liberal Party should implement policies that benefit liberal voters (in the hopes of deriving some votes from the New Democratic Party voter base) or conservative voters (in the hopes of deriving some votes from the Conservative Party voter base). The first option is likely easier because there is significant overlap between the demographics of the Liberal Party and the New Democratic Party allowing Liberal Party officials to serve their constituents as well as gain a win with larger margins in the next election. Implementing policies that will appease conservative voters will likely contradict liberal platforms but it is possible to implement policies that satisfy conservative demographics (improving pension plans for older Canadians, offering publicly funded trade schools for Canadians who have not attended postsecondary education, and so forth). Similar reasoning can be applied to consider how Conservative Party officials can enact policies to win the most seats in the 2023 Canadian Federal Election.

When comparing our popular vote prediction assuming 100% voter turnout to the results of the 2019 Canadian Federal Election, it appears that the Conservative Party did a better job than the Liberal Party at mobilizing its voter base to cast their ballots. In fact, our results suggest that if the Liberal Party was able to motivate more of their voter base to turn out to vote, they likely would have won the popular vote and would not need to worry about governing with a Liberal minority government (the number of seats won by other parties is greater than the number of seats won by the Liberal Party). In 2023, given the information gained from our analysis, the Liberal Party could implement initiatives to encourage greater turnout of their base by targeting demographics such as Canadian university students (young and highly educated). This could include measures like social media campaigns spreading awareness of their platforms or offering free rides to polling places from university campuses.

Although our confidence in our prediction of the popular vote has been limited for the reasons mentioned above, in the future, there are several ways we can increase the accuracy of our predictions on the individual level and on the population level. If we can find a survey and post-stratification dataset with detailed economic and ethnicity data, we can test other model implementations to determine if we can improve the accuracy of our multinomial logistic regression model. It might also be useful to find datasets that better illustrate the differences between Liberal Party and New Democratic Party voters or between Conservative Party and People's Party voters to improve the accuracy of our model. Additionally, it might be too ambitious to assume the same mathematical model can capture voting intention across Canada. It might be conducive to develop a separate model for every province and predict popular vote in that way. Lastly, a more comprehensive post-stratification dataset would have likely improved the accuracy of our popular vote prediction. Additionally, if the post-stratification dataset we find in the future has information about the demographics for each federal electoral district, this would enable us to predict the outcome of the Canadian Federal Election in addition to the popular vote.

# Appendix

## Post-stratification dataset collection cost
Statistics Canada did not disclose the true cost of conducting the survey but we can make some speculations based on the available information about their field work methodology. Surveying was conducted using Computer Assisted Telephone Interviewing (CATI) wherein interviewers read aloud the computerized questionnaire and immediately record the respondent's answers. Although this allows for a reduction in costs compared to traditional in-person surveying, labor costs still include time spent computerizing the survey, training interviewers, and having interviewers administer the questionnaire. Other labor costs include designing the questionnaire and surveying methodology as well as conducting quality control (data consistency was checked by the CATI system during surveying and unresolved inconsistencies were handled afterwards by support staff). Non-labor costs likely included paying for equipment, phone service, offices, and so forth. Again, although we don't have exact costs, we can conclude that the time and costs associated with conducting the GSS is a clear reason why it is only administered once a year.

## Derivation of Multinomial Logistic Regression Model
The multinomial logistic regression model consists of several binary logistic regression models. Like binary logistic regression, the multinomial logistic regression predicts the probability that the $i$th observation has outcome $k$ using the following function:

$$f(k, i) = \beta_{0,k} + \beta_{1,k}x_{1,i} + \beta_{2,k}x_{2,i} + ... + \beta_{M,k}x_{M,i}$$
where $\beta_{m,k}$ is the coefficient for the $m$th explanatory variable and the $k$th outcome while $x_{m, i}$ is the value of the $m$th explanatory variable for the $i$th observation. In our case, we have $M=5$ (age, sex, province, education, and religion importance) explanatory variables so the function as applicable to our model is:  

$$f(k, i) = \beta_{0,k} + \beta_{1,k}x_{1,i} + \beta_{2,k}x_{2,i} + \beta_{3,k}x_{3,i} + \beta_{4,k}x_{4,i} + \beta_{5,k}x_{5,i}$$
Note that we can represent $\beta_{0,k}, \beta_{1,k}, \beta_{2,k}, \beta_{3,k}, \beta_{4,k}, \beta_{5,k}$ and $1, x_{1,i}, x_{2,i}, x_{3,i}, x_{4,i}, x_{5,i}$  as row vectors $\boldsymbol{\beta_k}$ and $\boldsymbol{x_i}$, respectively. Then, the function can be simplified as follows:
$$f(k, i) = \boldsymbol{\beta_k} \cdot \boldsymbol{x_i}$$
where we take the dot product of the two row vectors we just defined.

As previously mentioned, the multinomial logistic regression model is a series of binary logistic regressions where the probability of each outcome of the response variable (vote choice for the 2019 Canadian Federal Election) is regressed against a chosen pivot outcome. Let $Y_i$ represent the outcome of the response variable for the $i$th observation. We have a total of 5 possible outcomes ("New Democratic Party", "Green Party", "Liberal Party", "Conservative Party", "People's Party", and "Bloc Québécois"), represented as 1, 2, 3, 4, 5, and 6 respectively. Let's choose the "Conservative Party" (or 4) as the pivot. In mathematical notation, this is:
$$ \ln{\frac{\Pr(Y_i = 1)}{\Pr(Y_i = 4)}} = \boldsymbol{\beta_1 \cdot X_i} $$
$$ \ln{\frac{\Pr(Y_i = 2)}{\Pr(Y_i = 4)}} = \boldsymbol{\beta_2 \cdot X_i} $$
$$\ln{\frac{\Pr(Y_i = 3)}{\Pr(Y_i = 4)}} = \boldsymbol{\beta_3 \cdot X_i} $$
$$\ln{\frac{\Pr(Y_i = 5)}{\Pr(Y_i = 4)}} = \boldsymbol{\beta_5 \cdot X_i} $$
$$\ln{\frac{\Pr(Y_i = 6)}{\Pr(Y_i = 4)}} = \boldsymbol{\beta_6 \cdot X_i} $$

Then, we solve for the probabilities by exponentiating both sides:
$$ \Pr(Y_i = 1) = \Pr(Y_i = 4) \cdot e^{\boldsymbol{\beta_1 \cdot X_i}} $$
$$ \Pr(Y_i = 2) = \Pr(Y_i = 4) \cdot e^{\boldsymbol{\beta_2 \cdot X_i}} $$
$$ \Pr(Y_i = 3) = \Pr(Y_i = 4) \cdot e^{\boldsymbol{\beta_3 \cdot X_i}}$$
$$ \Pr(Y_i = 5) = \Pr(Y_i = 4) \cdot e^{\boldsymbol{\beta_5 \cdot X_i}}$$
$$ \Pr(Y_i = 6) = \Pr(Y_i = 4) \cdot e^{\boldsymbol{\beta_6 \cdot X_i}}$$

The probability of the pivot outcome can be calculated because we know that the probability of all outcomes must sum to 1:
$$ \Pr(Y_i = 4) = 1 - \left(\Pr(Y_i = 4) \cdot e^{\boldsymbol{\beta_1 \cdot X_i}} + \Pr(Y_i = 4) \cdot e^{\boldsymbol{\beta_2 \cdot X_i}} + \Pr(Y_i = 4) \cdot e^{\boldsymbol{\beta_3 \cdot X_i}} + \Pr(Y_i = 4) \cdot e^{\boldsymbol{\beta_5 \cdot X_i}} + \Pr(Y_i = 4) \cdot e^{\boldsymbol{\beta_6 \cdot X_i}}\right) $$
$$ 1 = \Pr(Y_i = 4) + \Pr(Y_i = 4) \cdot e^{\boldsymbol{\beta_1 \cdot X_i}} + \Pr(Y_i = 4) \cdot e^{\boldsymbol{\beta_2 \cdot X_i}} + \Pr(Y_i = 4) \cdot e^{\boldsymbol{\beta_3 \cdot X_i}} + \Pr(Y_i = 4) \cdot e^{\boldsymbol{\beta_5 \cdot X_i}} + \Pr(Y_i = 4) \cdot e^{\boldsymbol{\beta_6 \cdot X_i}} $$
$$ 1 = \Pr(Y_i = 4) \left(1 + e^{\boldsymbol{\beta_1 \cdot X_i}} + e^{\boldsymbol{\beta_2 \cdot X_i}} + e^{\boldsymbol{\beta_3 \cdot X_i}} + e^{\boldsymbol{\beta_5 \cdot X_i}} + + e^{\boldsymbol{\beta_6 \cdot X_i}} \right) $$
$$ 1 = \Pr(Y_i = 4) \left(1 + \sum_{k\in \{1,2,3,5,6\}} e^{\boldsymbol{\beta_k \cdot X_i}} \right) $$
$$ \Pr(Y_i = 4) = \frac{1}{1 + \sum_{k\in \{1,2,3,5,6\}} e^{\boldsymbol{\beta_k \cdot X_i}}} $$
Having the expression for $\Pr(Y_i = 4)$, we can represent the probabilities of the other outcomes as follows:
$$ \Pr(Y_i = 1) = \frac{e^{\boldsymbol{\beta_1 \cdot X_i}}}{1 + \sum_{k\in \{1,2,3,5,6\}} e^{\boldsymbol{\beta_k \cdot X_i}}} $$
$$ \Pr(Y_i = 2) = \frac{e^{\boldsymbol{\beta_2 \cdot X_i}}}{1 + \sum_{k\in \{1,2,3,5,6\}} e^{\boldsymbol{\beta_k \cdot X_i}}} $$
$$ \Pr(Y_i = 3) = \frac{e^{\boldsymbol{\beta_3 \cdot X_i}}}{1 + \sum_{k\in \{1,2,3,5,6\}} e^{\boldsymbol{\beta_k \cdot X_i}}} $$
$$ \Pr(Y_i = 4) = \frac{1}{1 + \sum_{k\in \{1,2,3,5,6\}} e^{\boldsymbol{\beta_k \cdot X_i}}}$$
$$ \Pr(Y_i = 5) = \frac{e^{\boldsymbol{\beta_5 \cdot X_i}}}{1 + \sum_{k\in \{1,2,3,5,6\}} e^{\boldsymbol{\beta_k \cdot X_i}}} $$
$$ \Pr(Y_i = 6) = \frac{e^{\boldsymbol{\beta_6 \cdot X_i}}}{1 + \sum_{k\in \{1,2,3,5,6\}} e^{\boldsymbol{\beta_k \cdot X_i}}} $$

# References